A Gentle Introduction into Structural Causal Models

### 2021-04-22 paper/draft.Rmd
draft stuff

>>>>

## Time in Causality

In causal research, we are interested in the relationship of our variables outside of a fixed time frame. 
We want to capture relationships that hold outside of an available dataset and results are intended to be generalizable to other time periods (other datasets).

A lot of research ignores the distinction between structural equation models(SEM) and structural causal models (SCM).
Essentially, the SCM builds on the SEM but there are some differences.
Firstly, the structural equation model specification stems from disciplines like economics and psychology and is inherently a parametric specification. 
The SCM on the other hand is a non-parametric specification.
Secondly, the SCM allows one to create causal graphs.^[Note that we can also create causal graphs with other models like a bayesian causal network model]

Nevertheless, both models also have a lot in common.

Structural Causal Models specify the relationship in functional forms.
This is very different to working with conditional probabilities. 
There are two popular schools of thought on how to specify causal mechanisms.

There are a number of reasons why people are interested in structural causal models.
We can specify exogenous and endogenous variables in our model specification. 
When all our variables are fully specified within our model specification (endogenous variables only) we have a deterministic SCM.

Note that while this is very unreasonable because we assume we can explain all exogenous error terms and further we assume that there are no unobserved variables. 
Henceforth, when applying these models, they are seldom fully deterministic.

# Common Queries in Causal Models
